---
title: "simpleRegress"
author: "Ana Mendez"
date: "April 17th 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
Based on the analysis of WEIGHT as explained by NECK, the residual standard error (RSE) is 43.91, and the adjusted R-squared value is 0.87.

## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 
```{r}
mHEADLEN <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(mHEADLEN)
summary.lm(mHEADLEN)
```

Based on the observations, the residual standard error is 67.81, and the adjusted R-squared is 0.69. When ranking the models from best to worst, the most effective model is the one where WEIGHT is explained by NECK, followed by the model where WEIGHT is explained by HEADLEN.

NECK RSE is 43.91
HEADLEN RSE is 67.81

```{r}
mHEADLEN <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(mHEADWTH)
summary.lm(mHEADWTH)
```

We can see that for this plot, the residual standard error is 76.41 and the adjusted R squared is 0.6064. 
Best to worst ordered by RSE    
 NECK RSE is 43.91     
 HEADLEN RSE is 67.81       
 HEADWTH RSE is 76.41 
 
```{r}
mLENGTH <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(mLENGTH)
summary.lm(mLENGTH)
```
 
 For this plot, we can see that the RSE is 61.83 and the adjusted R-squared is 0.7423.
 Best to worst ordered by RSE  
 
 NECK RSE is 43.91      
 LENGTH RSE is 61.83     
 HEADLEN RSE is 67.81         
 HEADWTH RSE is 76.41
 
```{r}
mCHEST <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(mCHEST)
summary.lm(mCHEST)
```
For this last plot, we see that the residual standard error is 33.08, and the adjusted R-squared is 0.962. 
 

 